### 2016-05-28 sec7_optimization.tex
Added intro part

To formally describe the idea, we define the a \emph{candidate sequence} as follows:

@@

To reduce a sequence $T$ to a candidate sequence, we need to strip out its 
maximal $G$-connected subsequences which does not form to $K,L$. Such a reduction
takes two rounds scan of $T$ as shown in Algorithm~\ref{algo:simp_prune}. In the 
first round, the consecutive portions of $T$ with size less than $L$ are removed.
In the second round, the maximal $G$-connected sequences of size less than $K$ are
removed. Clearly the simplification algorithm runs in $O(|T|)$ time.
\begin{algorithm}
\caption{Edge Simplification}
\label{algo:simp_prune}
\begin{algorithmic}[1]
\Require $T$
\State{---Remove the consecutive portion with size less than $L$---}
\State $c \gets 0$
\For {$i \in (0,...,|T|)$}
	\If{$T[i] - T[i-1] != 1$} 
		\If{$i - c < L$} 
			\State $T$ remove $[c:i)$
		\EndIf
		\State $c \gets i$
	\EndIf
\EndFor
\State{---Remove the pseduo-consecutive portion with size less than $K$---}
\State $s\gets 1$, $c\gets 0$
\For{$i \in (0: |T|)$}
	\If{$T[i] - T[i-1] > G$}
		\If{$s < K$}   
			\State $T$ remove $[c:i)$
		\EndIf
		\State {$c \gets i$, $s \gets 1$}
	\Else
		\State $s++$
	\EndIf
\EndFor
\end{algorithmic}
\end{algorithm}

@@

\subsubsection{Forward closure checking}

\begin{example}
We use Figure~\ref{fig:star_partition} (c) to demonstrate the candidate pruning. As shown, at the initial stage, $\{3,6:1,4\}$ is pruned, since $\{1,4\}$ is not a candidate sequence. By temporal monotonicity, candidates containing objects $\{3,6\}$ can all be pruned. Therefore, we are able to directly prune $\{3,4,6\}$, $\{3,5,6\}$ and $\{3,4,5,6\}$.
\end{example}
%
%With the help of the \emph{Monotonic Property}, the number of new candidates in each level is greatly reduced. We verify this in the experiment session as well.
\subsection{Load Balancing}

\subsection{Duplication Detection}

\subsection{Handling Overlapping Clusters}

>>>>

To formally describe the idea, we define the \emph{candidate sequence} as follows:

@@

To reduce a sequence $T$ to a candidate sequence, we need to strip out its 
maximal $G$-connected subsequences which does not form to $K,L$. Such a reduction
takes two rounds scan of $T$ as shown in Algorithm~\ref{algo:simp_prune}. In the 
first round, the consecutive portions of $T$ with size less than $L$ are removed.
In the second round, the maximal $G$-connected sequences of size less than $K$ are
removed. Clearly the simplification algorithm runs in linear time.
\begin{algorithm}
\caption{Edge Simplification}
\label{algo:simp_prune}
\begin{algorithmic}[1]
\Require $T$
\State{---Remove the consecutive segment with size less than $L$---}
\State $c \gets 0$
\For {$i \in (0,...,|T|)$}
	\If{$T[i] - T[i-1] != 1$} 
		\If{$i - c < L$} 
			\State $T$ remove $[c:i)$
		\EndIf
		\State $c \gets i$
	\EndIf
\EndFor
\State{---Remove the $G$-connected segment with size less than $K$---}
\State $s\gets 1$, $c\gets 0$
\For{$i \in (0: |T|)$}
	\If{$T[i] - T[i-1] > G$}
		\If{$s < K$}   
			\State $T$ remove $[c:i)$
		\EndIf
		\State {$c \gets i$, $s \gets 1$}
	\Else
		\State $s++$
	\EndIf
\EndFor
\end{algorithmic}
\end{algorithm}

@@

\subsubsection{Forward closure checking}
Although leveraging \emph{temporal monotonicity} could largely prune false candidates and reduce
the apriori search space, it is ineffective when a \textit{true} pattern exists. 
For example, if a final pattern of a star $Sr_s$ is the \textit{union of all vertices} in the star, 
then in apriori, ${|Sr_s|}\choose{i + 1}$ candidates needs to be generated at 
each level $i$. This results in an exponential search space while
the output only contains one pattern.  
In general, when candidates at level $i$ collectively forms a true pattern, 
running aprior produces many wasted candidates.

Let $Lv_i$ be the set of candidates at level $i$ of Algorithm~\ref{algo:apriori_mining},
we use the \emph{forward closure} $FC_i$ to denote the union of the objects in
all candidates in $Lv_i$. Then, the \emph{forward closure checking} is stated as follows:
\begin{theorem}[Forward Closure Checking Rule]
Let $Lv_i$ be the candidates generated at level $i$ in Algorithm~\ref{algo:apriori_mining},
if $FC_i$ is a proper pattern, then it is safe to terminate Algorithm~\ref{algo:apriori_mining}
and directly output $FC_i$.
\end{theorem}

\begin{proof}
We prove by contradiction. Suppose there exists another pattern $P$ such that $P.O \neq FC_i$, let 
$X=P.O - FC_i \neq \emptyset$. Consider a subset of $P$ which contains $X$ with size $i+1$, (i.e., $P_1 \subseteq P, P_1\subseteq X, |P_1|=i+1$). Since $P$ is a proper pattern, then $P_1$ is also a proper pattern. Therefore $P_1 \in Lv_i$.
Then it follows $X$ is in the forward closure of $FC_i$, (i.e.,$X \in FC_i$), which contradicts with $X\notin FC_i$.
\end{proof}

It is notable that, as the level grows in Algorithm~\ref{algo:apriori_mining}, the closure $FC$
reduces, thus the pruning power of $FC$ would be stronger.

%To facilitate efficient mining, we adapt the \emph{forward closure checking} rule to 
%quickly determine whether it is sufficient to terminate the apriori search. Given the candidates at
%level $i$ $Lv_i$, we define the \emph{forward closure} of $Lv_i$ as the union of objects from
%all candidates. Then, the following \emph{forward closure checking} rule can be used:
%
%\begin{theorem}[Forward Closure Checking Rule]
%Let $Lv_i$ be the candidates generated at level $i$ in Algorithm~\ref{algo:apriori_mining}, the forward closure
%is $\mathbb{F}_i = \cup_{cand \in Lv_i} (cand.O) $. Algorithm~\ref{algo:apriori_mining} terminates when 
%$FC_i$ forms a valid GCMP .
%\end{theorem}

\begin{example}
We use Figure~\ref{fig:star_partition} (c) to 
demonstrate the power of candidate pruning. 
As shown, at the initial stage, $\{3,6:3\}$ is first pruned by \textit{Edge Simplification} since its
timestamps fails to be a candidate sequence. Subsequently, all further candidates containing $\{3,6\}$ 
are pruned by \textit{Temporal Monotonicity}. Then, we check the \textit{Forward Closure} 
of remaining candidates (i.e., $\{3,4\}$ and $\{3,5\}$) and find $\{3,4,5\}$ is a
valid candidate. Therefore, $\{3,4\}$ and $\{3,5\}$ are pruned, and $\{3,4,5\}$ is the output.
\end{example}
%
%With the help of the \emph{Monotonic Property}, the number of new candidates in each level is greatly reduced. We verify this in the experiment session as well.

\subsection{Optimal Star Partition and Load Balance}
As we see in Section~\ref{sec:spm_solution}, the apriori phase in SPM may
take exponential time. An important factor affecting the performance of
SPM is the size of stars. We notice that, the
size of star is related to the way the vertexes are numbered.

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{star-alt.eps}
\caption{An alternative numbering and partitioning of the connection graph in Figure~\ref{fig:star_partition}.}
\label{fig:star-alt}
\end{figure}

For example,
Figure~\ref{fig:star-alt} gives an alternative numbering of vertexes 
in connection graph thus produces a different set of stars. This partitioning constructs
four stars with the maximum star consisting of 5 edges. Compared to the partitioning
in Figure~\ref{fig:star_partition}(b) where five stars are produced with maximum star of 3 edges, this
partition is inferior in two aspects. First, this partition lacks of one star, which
results in a smaller number of parallelism. Second, this partition has a larger size
of maximum edges. With the exponential time complexity in Apriori, the largest star takes
much more time to compute.

The example shows that different numbering of vertexes in connection graph indeed affects
the SPM performance. Therefore we wish to find a numbering scheme of connection graph
such that the edges in each star are best distributed. Notice that the sum of edges 
is invariant for any numbering scheme of the graph. Thus, our goal is equivalent to find
a numbering such that the maximum edge size from all stars is minimized.

To formalize the objective, we design an linear algebra model as follows: Let $G$ denote a connection graph.
Let $\mathbb{A}$ be an arbitrary numbering of vertexes in $G$.
Let $(A:a_{i,j})$ be the matrix representing the induced graph wrt. $\mathbb{A}$. Let vector $\vec{b}$
be the \textit{one}\footnote{Every element in $\vec{b}$ is $1$} vector. Let $\vec{c}:c_j$ be the
vector equals to $A\vec{b}$. It is easy to see that each $c_j$ denotes the size of star for vertex $j$.
Therefore, the objective of load balancing can be formalized as follows:
\begin{equation}
\mathbb{A}  = \argmin(||A\vec{b}||_\infty) \text{,where } ||A\vec{b}||_\infty = \max_{1\leq j \leq n}(c_j)
\end{equation}

Though the objective is well-defined as above, it is challenging to directly optimize
the equation. First, suppose there are $n$ vertex in $G$, enumerating
all possible $\mathbb{A}$s leads to $n!$ combinations. 
Such a high complexity is trivially unpractical. Second,
since $G$ is constructed at runtime, 
the load planning can only start after $G$ is created. 
The planning time are thus required to be short enough otherwise the benefit cannot payoff the planning time.

Despite these challenges, we observe that there is a 
$O(1)$ time solution which is good enough as stated in the 
following theorem.

\begin{theorem}[Balance of Star Partition]
Let $G$ be a connection graph with degree $d$ and $n$ vertexes.
Let $\mathbb{A}^*$ be the optimal numbering wrt. Equation 1.
For any numbering, $\mathbb{A}$, with high probability, the 
absolute difference between $\mathbb{A}^*$ and $\mathbb{A}$ is $O(\sqrt{n \log n})$.
That is, it is very likely that 
$||\mathbb{A}\vec{b}||_\infty = ||\mathbb{A}^*\vec{b}||_\infty + O(\sqrt{n \log n})$.
\end{theorem}

\begin{proof}
Let $\mathbb{A}^*$ be the optimal solution wrt Equation 1. Since we have a star
for each object, by the pigeon-hole theorem, $||A^*\vec{b}||_\infty \geq d/2$.
Next, let $e_{i,j}$ be a 0-1 indicator variable determining whether vertex $i$ connects vertex $j$
in $G$. Note that edges in $G$ are independent. We use $d_i$ 
to denote the degree of object $i$ in $G$.  It follows that $E[d_i]=E[\Sigma_{1\leq j \leq n}e_{i,j}]=d$.
Since in the star partition, each edge is assigned to the ending vertex
with lower ID, the connection between $a_{i,j}$ and $e_{i,j}$ can be written as:
\begin{equation*}
a_{i,j} = \begin{cases}
			e_{i,j}, i>j \\
			0, otherwise
		  \end{cases}  
\end{equation*}
There are two observations made on the above equation. First, since $e_{i,j}$s are independent,
$a_{i,j}$s are independent. Second, since $i>j$ and $e_{i,j}$ are independent. 
$E[a_{i,j}] = E[e_{i,j}]E[i>j]= E[e_{i,j}]/2$.

By definition, $c_i = \Sigma_{1\leq j \leq n} a_{i,j}$, 
is a sum of $n$ independent 0-1 variables. Taking expectation on both sides, 
we get: $E[c_i] = E[\Sigma_{1\leq j \leq n} a_{i,j}]=E[\Sigma_{1\leq j \leq n} e_{i,j}]/2 = d/2$. Let $\mu =E[c_i] = d/2$, 
$t = \sqrt{n\log n}$, by Hoeffding's Inequality, the following holds:
\begin{equation*}
\begin{split}
	Pr(c_i \geq \mu + t) 
						&\leq \exp(\frac{-2t^2}{n}) \\
						&= \exp(-2\log n) = n^{-2}
\end{split}
\end{equation*}

The first step is due to the fact that all $a_{i,j}$ are bounded in the range of [0,1]. 
Next, since the event $(\max_{1\leq j \leq n}(c_j) > \mu + t)$ can be viewed as
$\cup_{c_i} (c_i > \mu + t )$, by Union Bound, we achieve the following:
\begin{equation*}
\begin{split}
	Pr(||A\vec{b}||_\infty > \mu + t) &=Pr(\max_{1\leq j \leq n}(c_j) > \mu + t)  \\
		&\leq \Sigma_{1 \leq i \leq n} Pr(c_i \geq \mu + t) \\
		& = n^{-1} = 1/n
\end{split}
\end{equation*}
Substitute $t$ and $\mu$, we achieve the following concise form:
\begin{equation*}
	Pr(||A\vec{b}||_\infty > (d/2 + \sqrt{n\log n})) \leq 1/n
\end{equation*}
This indicates that, the probability of $(||A\vec{b}||_\infty-d/2)$ being less than or equal to $ O(\sqrt{n\log n})$ is $(1-1/n)$. With the observed fact that $||A^*\vec{b}||_\infty \geq d/2$, we conclude that
with probability greater than $(1-1/n)$, 
the difference between $||A\vec{b}||_\infty$ and $||A^*\vec{b}||_\infty$ is less than $O(\sqrt{n\log n})$.
\end{proof}

In fact, we have a tighter bound of $||A\vec{b}||_\infty -||A^*\vec{b}||_\infty$ if 
the connection graph is \emph{dense}. Specifically, if $d\geq \sqrt{12\log n}$, the following
equation holds:
\begin{equation*}
Pr(||A\vec{b}||_\infty > (d/2 + O(\sqrt{d\log n})) \leq 1/n
\end{equation*}

Induced by the Theorem 5, taking object IDs as the numbering (i.e., in Algorithm~\ref{algo:spm_overview})
would be guaranteed to produce a partition which is almost optimal.
%By definition, $c_i = \Sigma_{1\leq j \leq n} a_{i,j}$, is a sum of $n$ independent 0-1 variables.
%Taking expectation on both sides, we get: $E[c_i] = E[\Sigma_{1\leq j \leq n} a_{i,j}]=E[\Sigma_{1\leq j \leq n} e_{i,j}]/2 = d/2$. Let $\mu =E[c_i] = d/2$, $\delta = \sqrt{12\log n/d}$, by Chernoff Bound, 
%the following holds:
%\begin{equation*}
%\begin{split}
%	Pr(c_i \geq (1+ \delta)\mu) & \leq \exp(-\mu \delta^2 / 3) \\
%						&= \exp(-d/2 \times 12 \log n / 3d ) \\
%						&= \exp(-2\log n) = n^{-2}
%\end{split}
%\end{equation*}
%
%Next, since the event $\max_{1\leq j \leq n}(c_j) > (1+ \delta)\mu$ can be viewed as
%$\cup_{c_i} (c_i > (1+ \delta)\mu )$, by Union Bound, we achieve the following:
%\begin{equation*}
%\begin{split}
%	Pr(||A\vec{b}||_\infty > (1+ \delta)\mu)) &=Pr(\max_{1\leq j \leq n}(c_j) > (1+ \delta)\mu)  \\
%		&\leq \Sigma_{1 \leq i \leq n} Pr(c_i \geq (1+ \delta)\mu) \\
%		& = n^{-1} = 1/n
%\end{split}
%\end{equation*}
%Substitute $\mu$ and $\delta$, we achieve the following concise form:
%\begin{equation*}
%	Pr(||A\vec{b}||_\infty > (d/2 + \sqrt{3 d\log n})) \leq 1/n
%\end{equation*}
%This indicates that, the probability of $(||A\vec{b}||_\infty-d/2)$ being less than $ O(\sqrt{d\log n})$ is $(1-1/n)$. Notice that $||A^*\vec{b}||_\infty \geq d/2$, therefore, with probability greater than $(1-1/n)$, the difference between $||A\vec{b}||_\infty$ and $||A^*\vec{b}||_\infty$ is less than $O(\sqrt{d\log n})$.
%\end{proof} 
%
%Theorem 5 is powerful enough to induce the following lemma:
%\begin{theorem}[Load Balance of Algorithm~\ref{algo:spm_overview}]
%The partitioning in Algorithm~\ref{algo:spm_overview} differs at most $O(\sqrt{d\log |\mathbb{O}|})$ 
%from optimal partitioning with probability $(1-1/|\mathbb{O}|)$.
%\end{theorem}
